Monitoring of Onion Maggot (\textit{Delia antiqua}) fly through improved trapping

### 2019-08-14 manuscript/journal-applied-entomology/main.tex
added introduction

\title{Camila Title Suggestions?}

% List abbreviations here, if any. Please note that it is preferred that abbreviations be defined at the first instance they appear in the text, rather than creating an abbreviations list.
\abbrevs{ABC, a black cat; DEF, doesn't ever fret; GHI, goes home immediately.}

% Include full author names and degrees, when required by the journal.
% Use the \authfn to add symbols for additional footnotes and present addresses, if any. Usually start with 1 for notes about author contributions; then continuing with 2 etc if any author has a different present address.
\author[1\authfn{1}]{Author One PhD}
\author[2\authfn{1}]{Author A.~Two MD}
\author[2\authfn{2}]{Author Three PhD}
\author[2]{Author B.~Four}

@@

% Include full affiliation details for all authors
\affil[1]{Department, Institution, City, State or Province, Postal Code, Country}
\affil[2]{Department, Institution, City, State or Province, Postal Code, Country}

\corraddress{Author One PhD, Department, Institution, City, State or Province, Postal Code, Country}
\corremail{correspondingauthor@email.com}

\presentadd[\authfn{2}]{Department, Institution, City, State or Province, Postal Code, Country}

\fundinginfo{Funder One, Funder One Department, Grant/Award Number: 123456, 123457 and 123458; Funder Two, Funder Two Department, Grant/Award Number: 123459}

% Include the name of the author that should appear in the running header
\runningauthor{Author One et al.}

@@

\section{Introduction}
Coming Soon...

@@

To improve trap catch and monitoring of adult onion maggot fly, iterations of sticky traps of various shape, size, and color were evaluated in 2005 and 2006 in Upstate New York.  Additionally, after an improved shape, size, and color trap was developed, this trap was tested in conjunction with a Delia Lure attractant to evaluate potential improvement in trap catch.

All shape, size, and color trials were conducted in commercial fields on muck soil near Elba, NY in western upstate New York (\textbf{GPS coordinates, other information about unique muck characteristics?}).  Trials evaluating trap catch in conjunction with Delia Lure were evaluated both in Elba, NY and on muck soil near Potter, NY in western upstate New York.  No modifications were done to commercial weed, disease, or pest management and spray programs (which closely followed recommended management \citep{reiners2011integrated}), the primary target of which is \textit{Thrips tabaci}.  This management is unlikely to affect \textit{D. antiqua} populations due to spray timing and minimal impacts of thrips sprays on onion maggot fly \citep{finch1986behavior}.

@@

\paragraph{Shape Trial} To evaluate effect of trap shape on trap catch and monitoring of adult onion maggot fly, square, cube, short cylindrical, tall cylindrical, and spherical semi-gloss white shapes each with a surface area of 180 $cm^2$  were coated with sticky material (\textbf{Tanglefoot?}) and placed along the long edge of an onion (\textit{Allium cepa} L.) field near Elba, NY.  Traps were spaced 15.24 meters apart and arranged in a randomized control block design with five replications.  Traps were placed in the field in late May then collected in late June (\textbf{Not sure about this, more info needed?}).

@@

\printendnotes

@@

>>>>

\title{Monitoring of Onion Maggot (\textit{Delia antiqua}) fly through improved trapping}

% List abbreviations here, if any. Please note that it is preferred that abbreviations be defined at the first instance they appear in the text, rather than creating an abbreviations list.
%\abbrevs{ABC, a black cat; DEF, doesn't ever fret; GHI, goes home immediately.}

% Include full author names and degrees, when required by the journal.
% Use the \authfn to add symbols for additional footnotes and present addresses, if any. Usually start with 1 for notes about author contributions; then continuing with 2 etc if any author has a different present address.
\author[1\authfn{1}]{Denis S. Willett}
\author[1\authfn{1}]{Camila C. Filgueiras}
\author[2]{Starker Wright}
\author[1]{Jan P. Nyrop}
\author[1]{Brian A. Nault}

@@

% Include full affiliation details for all authors
\affil[1]{Department of Entomology, Cornell AgriTech, Cornell University, Geneva, NY, 14456, USA}
\affil[2]{Bartlett Tree Experts, Dublin, PA, USA}

\corraddress{Denis S. Willett, 15 Castle Creek Drive, Geneva, NY 14456}
\corremail{deniswillett@cornell.edu}

% Include the name of the author that should appear in the running header
\runningauthor{Willett et al.}

@@

\section{Introduction}

The onion maggot (\textit{Delia antiqua}) is a economically important pest of Allium crops worldwide.  Found primarily in temperate climates, \textit{D. antiqua} is well established in onion growing regions in the Americas, Europe and Asia where crop losses due to this pest can range from 50 to 100 percent on Onions (\textit{Allium cepa} L.), Garlic (\textit{Allium sativem} L.), Scallions (\textit{Allium fistulosum} L.), and Chives (\textit{Allium schoenoprasum} L.) if left unmanaged \citep{ellis1979factors,ning2017predicting,nault2007ecology, nault2006performance, nault2006onion}.  Intensive management of this pest has become a necessary component of effective onion production in many parts of the world and severe damage from this pest continues to occur in areas where Allium crops are grown without rotation or where resistant populations have developed \citep{martinson1988dispersal, nault2006onion}.

Onion maggot is a important pest in part due to its lifecycle.  It is multivoltine with three generations per year in the Northern United States \citep{eckenrode1975population, hoepting2004insecticide}.  Larvae feed on plant roots and allium bulbs causing the damage associated with this pest.  First generation infestations are particularly devastating as early feeding by onion maggot larvae can kill seedlings \citep{nault2006onion, nault2006performance}.  In addition, damaged plants become more susceptible to infestation by subsequent generations, pathogens, and damage at harvest time \citep{eckenrode1986impact,nault2006performance}.  Even if plants are not completely killed by onion maggot larvae, feeding damage often renders the produce unmarketable.

Management of this pest has relied on in-furrow insecticide application at planting \citep{nault2006performance}.  A variety of organophosphate and carbamate insecticides have been used and discontinued \citep{nault2006performance}.  Chlorpyrifos is commonly used as a drench treatment currently in combination with a cyromazine seed treatment \citep{nault2006performance}.  Fipronil and spinosad are being considered as effective alternatives as some onion maggot populations are developing resistance to chlorpyrifos \citep{nault2006onion}.  Cultural practices such as rotation, removal of damaged onions, and delayed planting remain effective management alternatives \citep{martinson1988dispersal, finch1985influence, nault2011delaying}.

Delayed planting can be an effective control tactic because onion development if offset phenologically from development of onion maggot oviposition and larval development \citep{nault2011delaying}.  Such tactics, and effective control in general, rely on monitoring onion maggot populations in the field.  While useful in areas where onion maggot populations are already established, effective monitoring becomes even more important where onion maggot is a relatively recent introduction or in areas where producers are not sure they have a problem.

Past efforts at monitoring and understanding onion maggot population dynamics have relied on three principle methods: sticky traps, water traps, and simulation.  Sticky traps and water traps have been used to varying degrees of success in North America and Europe but often suffer from limitations related to inefficiencies in counting \citep{thomingdeveloping,otto2000development}.  Trap catch of adult onion maggot flies can be low compared to field populations.  Additionally, the number of onion maggots caught compared to other insect collected in traps necessitates a great deal of time in evaluating trap catches.  Odor sources of attractants have also been added to traps in an effort to make them more attractive, but with limited success.  Simulation has relied on field data from traps, phenological studies in the lab, and climate information \citep{thomingdeveloping,otto2000development,ning2017predicting}.  Population simulations can prove useful at larger scales, but often lack extremely detailed location specific information necessary for producers to determine if they have a severe problem in a specific field.

To further develop effective monitoring methodologies for the onion maggot, we evaluated a number of factors likely to influence attraction of adult onion maggot flies.  Previous work has suggested that color, shape, size, and chemistry all affect attraction of adult onion maggots \citep{harris1983color,harris1988host, thomingdeveloping, otto2000development}. We expanded on previous work in field trials over two years in upstate New York where a variety of novel trap designs with varying shape, size, color, and chemistry were presented to populations of adult onion maggot flies to evaluate monitoring potential and efficacy.

@@

To improve trap catch and monitoring of adult onion maggot fly, iterations of sticky traps of various shape, size, and color were evaluated in 2005 and 2006 in upstate New York.  Additionally, after an improved shape, size, and color trap was developed, this trap was tested in conjunction with a Delia Lure attractant to evaluate potential improvement in trap catch.

All shape, size, and color trials were conducted in commercial dry bulb onion fields on muck soil near Elba, NY in western upstate New York. Muck soils are high-organic matter soils formed from old lake beds.  Trials evaluating trap catch in conjunction with Delia Lure were evaluated both in Elba, NY and on muck soil near Potter, NY in western upstate New York.  All locations used in these studies are characterized by continuous onion production with no-rotation and consistently high onion maggot populations.  No modifications were done to commercial weed, disease, or pest management and spray programs (which closely followed recommended management \citep{reiners2011integrated}), the primary target of which is \textit{Thrips tabaci}.  This management is unlikely to affect \textit{D. antiqua} populations due to spray timing and minimal impacts of thrips sprays on onion maggot fly \citep{finch1986behavior}.

@@

\paragraph{Shape Trial} To evaluate effect of trap shape on trap catch and monitoring of adult onion maggot fly, square, cube, short cylindrical, tall cylindrical, and spherical semi-gloss white shapes each with a surface area of 180 $cm^2$  were coated with sticky material (\textbf{Tanglefoot?}) and placed along the long edge of an onion (\textit{Allium cepa} L.) field near Elba, NY.  Traps were spaced 15.24 meters apart and arranged in a randomized control block design with five replications.  Traps were placed in the field in late May then collected in late June.

@@

%\printendnotes

@@

\section{References}

\bibliographystyle{vancouver-authoryear}
\bibliography{om-attraction}
